---
title: "p8105_homework_1"
output: github_document
author: "Darwin Keung"
date: 2018-09-19
---

## Homework 1

### Problem 1


```{r setup, include=FALSE}
rm(list = ls())
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

This problem focuses on variable types and coercion.

The chunks below creates a data frame comprised of:

* A random sample of size 10 from a uniform[0, 5] distribution

* A logical vector indicating whether elements of the sample are greater than 2

* A (length-10) character vector

* A (length-10) factor vector

```{r problem_1_data frame, collapse=TRUE}
## Creating the data frame. ## `runif` generates random uniformly distributed numbers
variable_df = tibble(
  vec_random = runif(10 , min = 0, max = 5),
  vec_logical = vec_random > 2,
  vec_char = c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J"),
  vec_factor = factor(c("h", "e", "l", "l", "o", "w", "o", "r", "l", "d"))
)
## Taking the means of each variable in dataframe
mean(variable_df$vec_random)
mean(variable_df$vec_logical)
mean(variable_df$vec_char)
mean(variable_df$vec_factor)
```
The variable `vec_random` has class `r class(variable_df$vec_random)` and the variable `vec_logical` has class `r class(variable_df$vec_logical)`. `mean` works on numeric and logical vectors since there are quantities to use for the mean.

The variable `vec_char` has class  `r class(variable_df$vec_char)` and the variable `vec_factor` has class  `r class(variable_df$vec_factor)`. `mean` does not work on character or factor vectors - there are no numerical values to serve as input.

```{r as.numeric function, eval = FALSE}
## Coercing our variables into numeric. 
as.numeric(variable_df$vec_logical)
as.numeric(variable_df$vec_char)
as.numeric(variable_df$vec_factor)
```

The logical and factor variable can be converted to numeric. But character variable can't be converted to numeric.

```{r as.numeric conversions}
## convert character variable from character to factor to numeric; 
vec_char_to_factor = as.factor(variable_df$vec_char)
as.numeric(vec_char_to_factor)

## convert factor variable from factor to character to numeric. 
vec_fact_to_char = as.character(variable_df$vec_factor)
as.numeric(vec_fact_to_char)
```
Converting characters to factors assigns each character a catergorical value, which then can be converted from a factor value to a numeric value. 
Going from factor to character removes the catergorical value (e.g. A = 1) and turns it into a character (A = "A") which no longer can be turned into a numeric value.

### Problem 2

This problem focuses on plotting and the use of inline R code.

The chunk below creates a data frame comprised of:

* x: a random sample of size 1000 from a standard Normal distribution

* y: a random sample of size 1000 from a standard Normal distribution

* A logical vector indicating whether the x + y > 0

* A numeric vector created by coercing the above logical vector

* A factor vector created by coercing the above logical vector

``` {r Problem_2_Dataframe}
p2_df = tibble(
  x = rnorm(1000),
  y = rnorm(1000),
  sum = x + y > 0,
  num_sum = as.numeric(sum),
  fact_sum = as.factor(sum)
)

## should I name it sum or no name just use x + y > 0.
```

* The above dataframe, "p2_df" is `r dim(p2_df)`

* The mean is `r mean(p2_df$x)` and the median is `r median(p2_df$x)`

_?Do I need to round or trincate? How can I do this?_

* The proportion of cases for which the logical vector is `TRUE` is `r mean(p2_df$num_sum)*100`

_I'm not sure if I should use the sum function here then devide by the nrow(df). `mean` works if we are taking the proportion of 0 and 1 statements._

```{r scatterplot}
## scatterplot of x and y
ggplot(p2_df, aes(x = x, y = y)) + geom_point(aes(color = sum))
ggsave("hw1p2_scatterplot_xy.pdf", height = 4, width = 6)

## scatterplot using numeric
ggplot(p2_df, aes(x = x, y = y)) + geom_point(aes(color = num_sum))

## scatterplot using factor variables
ggplot(p2_df, aes(x = x, y = y)) + geom_point(aes(color = fact_sum))
```

*Comment on color scales*

The color scale of the scatterplot of x and y as well as the scatterplot using factor variables uses True and False and a more distinct color scheme. The numeric scatter plot color scale is a gradient assigning colors for intervals between 0 and 1, but we only have 0 and 1 and thus two colors shown.
The color scale of the scatterplot of x and y as well as the scatterplot using factor variables uses True and False and a more distinct color scheme. The numeric scatter plot color scale is a gradient assigning colors for intervals between 0 and 1, but we only have 0 and 1 and thus two colors shown.