---
title: "p8105_hw1_dk2759"
output: github_document
author: "Darwin Keung"
date: 2018-09-19
---

## Homework 1

### Problem 1

This problem focuses on variable types and coercion.

The chunk(s) below creates a data frame comprised of:
* A random sample of size 10 from a uniform[0, 5] distribution
* A logical vector indicating whether elements of the sample are greater than 2
* A (length-10) character vector
* A (length-10) factor vector

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```


```{r problem_1_data frame}
variable_df = tibble(
  vec_random = runif(10 , min = 0, max = 5),
## `runif` generates random uniformly distributed numbers, n is the number of observations. Min, Max are the lower and upper limits of the distribution.
  vec_logical = vec_random > 2,
  vec_char = c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J"),
  vec_factor = factor(c("h", "e", "l", "l", "o", "w", "o", "r", "l", "d"))
)
variable_df

## Taking the means of each variable in dataframe
mean(variable_df$vec_random)

mean(variable_df$vec_logical)

mean(variable_df$vec_char)

mean(variable_df$vec_factor)
```


The variable `vec_random` has class `r class(variable_df$vec_random)` and the variable `vec_logical` has class `r class(variable_df$vec_logical)`. `mean` works on numeric and logical vectors since there are quantities to use for the mean.

The variable `vec_char` has class  `r class(variable_df$vec_char)` and
the variable `vec_factor` has class  `r class(variable_df$vec_factor)`. `mean` does not work on character or factor vectors - there are no numerical values to serve as input.

```{r as.numeric function}

### Problem 2